

## Part III

Now practice with data from our class. Please create a **person-network** with the data set 

hudk4050-classes.csv

To create this network you will need to create a person-class matrix using the tidyr functions and then create a person-person matrix using `t()`. 
You will then need to plot a matrix rather than a data frame using igraph.

```{r}
df <- read.csv("HUDK4050-2018-classes.csv")
```

```{r}
library(dplyr)
library(tidyr)
```

```{r}
# slice/subset rows
df2 <- df[-c(1,2), ]
df2 <- slice(df, -(1:2))

# select columns
df3 <- select(df2, Q8:Q6)
df3 <- select(df2, Q8, Q9, Q1, Q3, Q4, Q5, Q6, Q7)
df3 <- df[,15:21]

# subset rows and columns
df4 <- df[-c(1,2), 15:22]

# change header / variable names
colnames(df4) <- c("First Name", "Last Name", "Class 1", "Class 2", "Class 3", "Class 4", "Class 5")
df4 <- rename(df4, "First Name" = Q8)
names(df4) <- unlist(df[1,15:22])

# clean spaces 
df4$`Class 1` <- gsub('\\s+', '', df4$`Class 1`)
df4$`Class 2` <- gsub('\\s+', '', df4$`Class 2`)
df4$`Class 3` <- gsub('\\s+', '', df4$`Class 3`)
df4$`Class 4` <- gsub('\\s+', '', df4$`Class 4`)
df4$`Class 5` <- gsub('\\s+', '', df4$`Class 5`)
df4$`Class 6` <- gsub('\\s+', '', df4$`Class 6`)

# change variable to character
df4$`Class 1` <- as.character(df4$`Class 1`)

# change dataframe to character dataframe
df4[] <- lapply(df4, as.character)

# this won't work on factors, must change to characters
df4$`Class 1` <- gsub(" ", "", df4$`Class 1`)
df4$`Class 2` <- gsub(" ", "", df4$`Class 2`)
df4$`Class 3` <- gsub(" ", "", df4$`Class 3`)
df4$`Class 4` <- gsub(" ", "", df4$`Class 4`)
df4$`Class 5` <- gsub(" ", "", df4$`Class 5`)
df4$`Class 6` <- gsub(" ", "", df4$`Class 6`)

# this didn't work during workshop
for (row in 1:ncol(df4)) {
  for (col in 3:nrow(df4)) {
    df[col,row] <- gsub('\\s+', '', df4[col,row])
  }
}

# combine first and last name
df5 <- unite(df4, "Name", "First Name", "Last Name")

# delete empty row
df6 <- filter(df5, Name != "_")

# create two variables: student_name, course_name
df7 <- gather(df6, key=class, n, 2:7, na.rm=TRUE) %>% select(-class)
df7 <- filter(df7, n > " ")

# remove duplicate students
df8 <- unique(df7)
```

# Goal: You will then need to plot a matrix rather than a data frame using igraph.

```{r}
# create person to class matrix
df8$value <- 1 # add a value column to spread by
df9 <- spread(df8, n, value, fill=0) 
df10 <- select(df9, -"Name") 
# %>% as.matrix()
row.names(df10) <- df9[,1]
df10 <- as.matrix(df10)
class(df10) # confirm matrix class
dim(df10) # confirm dimensions of matrix

# create person to person matrix using t()
df11 <- df10 %*% t(df10)

library(igraph)
g <- graph_from_adjacency_matrix(df11)
plot(g, layout=layout.fruchterman.reingold, edge.arrow.size=0)


```




